
% Thesis Abstract -----------------------------------------------------


%\begin{abstractslong}    %uncommenting this line, gives a different abstract heading
\begin{abstracts}        %this creates the heading for the abstract page

  %% - Neutrinos are important blah blah
  %% - km3net is a thing focused on researching neutrinos
  %% - event trigger and gpu pipeline do blah
  %% - goal of this paper
  %% - summarize RQs
  %% - summarize entire pipeline
  %% - summarize result
  %% - a line from conclusion + recommendation

  Neutrinos are highly elusive subatomic particles which can only be
  detected with the help of large neutrino detectors. The KM3NeT
  neutrino telescope is one such detector currently being constructed
  at the bottom on the Mediterranean Sea. Due to its large volume and
  the presence of background noise, ``event trigger'' algorithms are
  utilized by the data acquisition pipeline of the detector to sift
  through the noise. Although the state of the art event trigger
  algorithms are able to perform data filtration in real time, the
  quality of filtration is poor. A GPU Pipeline was developed to
  improve the quality of filtration without compromising the runtime
  performance. The goal of this paper is to improve upon this GPU
  Pipeline using Artificial Neural Networks. The paper explores the
  possibility of replacing parts of the GPU Pipeline using Multi Layer
  Perceptrons and Graph Convolusional Neural Networks. The results are
  promising and new avenues of research are discovered through this
  work.

  \emph{\textbf{Keywords}: Neutrino detection, Deep learning, Graph
  Neural Networks, Geometric Learning, KM3NeT.}
\end{abstracts}
%\end{abstractlongs}


% ---------------------------------------------------------------------- 
